---
output: html_document
---


```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
require("ggplot2")
require ("plyr")
require ("caTools")
```

## <font color="midnightblue">Employment Data Analysis</font> {.tabset .tabset-fade}

**The secrets to getting ahead in this company:**  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;1. Work for many years with several different companies before coming to this company. Your Total Time Worked will provide you with a higher job level and higher monthly income.  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;2. Change Managers frequently.  The longer you stay with a manager the longer the time before you will receive you next promotion.  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;3. Aspire and work toward being a manager or director.  These job roles make more money. If you stay with the company long enough and have many years of total work time before coming to the company, you should be able to obtain one of these position.  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;4. Do not work as a Sales Representative.  This job role makes less money and is most likely to quit.  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;5. Don't worry about job performance ratings.  Everyone receives either an outstanding or an excellent rating.  The ratings have not impact on your monthly income.  



###Executive Summary
####<font color="midnightblue">Executive Summary</font>
#####<font color="midnightblue">Introduction</font>
This project was created for DDSAnalytics, an analytics company that specializes in talent management solutions for Fortune 1000 companies. According to DDSAnalytics, "Talent management is defined as the iterative process of developing and retaining employees. It may include workforce planning, employee training programs, identifying high-potential employees and reducing/preventing voluntary employee turnover (attrition)." DDSAnalytics has hired our team of data scientists to predict employee turnover by analyzing existing employee data. The company can utilize this analysis to predict factors that lead to employee attrition and potentially prevent voluntary employee attrition.  

#####<font color="midnightblue">Project Goals</font>  
The main goal of this project is to analyze factors that lead to employee attrition to discover the top most influential factors that lead to attrition.  We are also interested in specifics about the company to help DDSAnalytics better understand its employees. We will analyze interesting trends in job roles at the current company as well as job satisfaction. This knowledge can ultimately help DDSAnalytics remain competitive when hiring employees and creating a positive company culture with satisfied employees.  

#####<font color="midnightblue">Findings</font>  
**Attrition Analysis:**  
Overall, we found that the number one cause of employee attrition is being a Sales Representative. Sales Representatives are 8 times more likely to have attrition than any other category. The second highest cause of attrition is working overtime; employees who work overtime are 5 times more likely to quit. Additionally, people who travel frequently are 4 times more likely to voluntarily leave the company. After these top three factors, the next most influential factors that lead to attrition are: being a Laboratory Technician, being in Human Resources, being a Sales Executive, travelling rarely, being a Research Scientist, being a manager, having a high number of companies worked, and living a far distance from the company.  
**Job Roles:**  
Since we are also interested in analyzing the current company, we ran the rest of our analysis only with current employees.  We found that the manager is the only job role that overlaps departments and that managers and directors tend to be older and have a higher monthly income than other roles (as expected). The only job role where males do not outnumber females in this company is the Manufacturing Director. The most interesting analysis conducted was on Sales Representatives, since this is the position that is most likely to quit. Sales Representatives have the lowest job levels, the lowest average duration at the company, and change managers the most frequently. All of these factors could contribute to their high turnover rates.  
**Company Specifics:**  
Again, we analyzed the company by only including current employees in our analysis. We found that 67% of the company works in Research and Development and that there are 46% more male employees than female employees. We also noticed that job level is strongly correlated to monthly income, which is based on work experience rather than performance. Since everyone in the company receives a performance rating of excellent or outstanding, this holds no weight on income or job satisfaction. Another interesting finding is that the longer employees work for the same manager, there tended to be a longer time there in between promotions. There is a relatively new group of employees at company with 50% of the company working there for less than or equal to 5 years.  
**Satisfaction:**  
Generally, job satisfaction increases as hourly rate decreases, suggesting that hourly rate could be a significant factor leading to satisfaction. Employees with PhD’s, Sales Representatives, and the HR Manager are the most satisfied, which is intriguing given Sales Representatives are the most likely to voluntarily quit. The most dissatisfied employees are the R&D Directors.  
**Interesting Findings:**  
There seems to be some discrepancies or errors in the data – we would expect a multiplicative relationship between daily rate and hourly rate or daily rate and monthly rate, but we do not see consistency in this data.  
**Experimentation:**  
For our attrition analysis, we ran a logistic regression and fit our model using the STEP procedure as our selection procedure. We ran our model against our test data and came out with 91.8% accuracy in our model, so we sorted the factors that lead to attrition by p-values in descending order of influence. For the remaining analysis on the company, we used linear regression modeling.  

#####<font color="midnightblue">Conclusion</font>
This analysis can help DDSAnalytics gain a competitive edge in this industry by creating a positive climate and culture and decreasing the attrition rate. With only 16% attrition, DDSAnalytics does not have a very high attrition rate to begin with. Our recommendations to lower this percentage even further are as follows:  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;1. Focus time and energy on your Sales Representatives. As this job role was the leading cause of attrition, it would be beneficial to ensure these employees remain satisfied. Things to consider altering about this job role could be the job level the frequent manager changes.  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;2. Do not grant as many overtime hours – employees are more likely to leave if they work overtime, so limit the number of employees that are working overtime hours.  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;3. Check in with R&D Directors – this is the most dissatisfied job role in your company.  

Please see our report for more detailed findings and experimentation. Please contact Noelle Brown, William Gonzalez, or Ann Nelson with any additional questions.  

###Data Set
####<font color="midnightblue">Data Set</font>
```{r include=TRUE, echo=FALSE}
# Read the datafile
attrit <- read.csv("Data/CaseStudy2-data.csv")
# Rearrange the columns to aid analysis

AttritColNames <- c( "Attrition", "BusinessTravel","Department","EducationField","Gender",  "JobRole", "Over18", "OverTime",  "MaritalStatus", "Age", "DailyRate","DistanceFromHome","Education",  "EmployeeCount", "EmployeeNumber", "EnvironmentSatisfaction",  "HourlyRate", "JobInvolvement", "JobLevel","JobSatisfaction","MonthlyIncome", "MonthlyRate", "NumCompaniesWorked",  "PercentSalaryHike", "PerformanceRating", "RelationshipSatisfaction", "StandardHours", "StockOptionLevel", "TotalWorkingYears", "TrainingTimesLastYear","WorkLifeBalance", "YearsAtCompany", "YearsInCurrentRole", "YearsSinceLastPromotion","YearsWithCurrManager")
attrit <- attrit[,AttritColNames]

#Data Set Description
cat("Number of employees in the dataset:")
cat(nrow(attrit))
cat("Number of variables in the dataset:")
cat(ncol(attrit))

cat ("Number of employees who have left the company and the number if employees who remain:")
print.data.frame(as.data.frame(table(attrit$Attrition,dnn=list("Left"))),row.names = FALSE)

# Eliminate variables that have no variation (constant values)
cat("3 Variables had no variation,that is, are constants and will not aid in the analysis")
cat("These variables are removed from the dataset.")

attrit$EmployeeCount <- NULL
attrit$StandardHours <- NULL
attrit$Over18 <- NULL

cat("Number of variables in the dataset, after removing constants:")
cat(ncol(attrit))

cat("Count of field types in the dataset:")
table(as.data.frame(sapply(attrit,class)))



```
#### <font color="midnightblue">Job Roles</font> 

```{r, fig.width=10}
ggplot(data=as.data.frame(table(attrit$Department,attrit$JobRole,dnn=list("Department","JobRole"))), aes(x= reorder(JobRole,Freq),y=Freq, fill = Department)) +
geom_bar(stat="identity") + coord_flip() + ggtitle("Number of Employees per Job Role by Department") + xlab("") + ylab("")+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4")))
```

##### The Job Role "Manager" is the only Job Role that is in multiple departments
##### To aid analysis the values of JobRole for the Managers will be recoded

```{r, fig.width=10}
# Recode Managers
attrit$JobRole <-with(attrit, ifelse(attrit$JobRole == "Manager",paste(attrit$Department,attrit$JobRole), as.character(attrit$JobRole)))
attrit$JobRole <- as.factor(attrit$JobRole)
ggplot(data=as.data.frame(table(attrit$Department,attrit$JobRole,dnn=list("Department","JobRole"))), aes(x= reorder(JobRole,Freq),y=Freq, fill = Department)) +
geom_bar(stat="identity") + coord_flip() + ggtitle("Number of Employees per Job Role by Department \nafter recoding") + xlab("") + ylab("")+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4")))

#Recode the variable "JobRole" to a numeric value for modeling
attrit$JRCode[attrit$JobRole=="Healthcare Representative"] <- 1L
attrit$JRCode[attrit$JobRole=="Human Resources"] <- 2L
attrit$JRCode[attrit$JobRole=="Laboratory Technician"] <- 3L
attrit$JRCode[attrit$JobRole=="Manufacturing Director"] <- 4L
attrit$JRCode[attrit$JobRole=="Research Director"] <- 5L
attrit$JRCode[attrit$JobRole=="Sales Executive"] <- 6L
attrit$JRCode[attrit$JobRole=="Research Scientist"] <- 7L
attrit$JRCode[attrit$JobRole=="Sales Representative"] <- 8L
attrit$JRCode[attrit$JobRole=="Human Resources Manager"] <- 9L
attrit$JRCode[attrit$JobRole=="Research & Development Manager"] <- 10L
0.0008

#Create a dataframe of current employees only
noattrit <- attrit[attrit$Attrition=="No",]
noattrit$Attrit <- NULL

```




### Attrition Analysis
#### <font color="midnightblue">Attrition</font> 
#### First take a quick look at the data. We will name the main dataset used for the attrition portion "attrition" 

```{r, include=TRUE, echo=TRUE}
attrition<-read.csv('Data/CaseStudy2reorder.csv', header = T)
#View(attrition)
head(attrition)
```

#### Drop the first Column, attach dataset for ease of variable recall 
```{r, include=TRUE, echo=TRUE, fig.width=10}
attrition<-attrition[,-1] # 1st col obsolete 
str(attrition)
summary(attrition)
```
#### Some of the categorical variables will not work in an analisys because they are not factors. Convert them to factors. 

```{r, include=TRUE, echo=TRUE, fig.width=10}
attrition$Education<-as.factor(attrition$Education) 
attrition$EnvironmentSatisfaction<-as.factor(attrition$EnvironmentSatisfaction)  
attrition$JobInvolvement<-as.factor(attrition$JobInvolvement)  
attrition$JobLevel<-as.factor(attrition$JobLevel)  
attrition$JobSatisfaction<-as.factor(attrition$JobSatisfaction)  
attrition$PerformanceRating<-as.factor(attrition$PerformanceRating)  
attrition$RelationshipSatisfaction<-as.factor(attrition$RelationshipSatisfaction)  
attrition$StockOptionLevel<-as.factor(attrition$StockOptionLevel)  
attrition$WorkLifeBalance<-as.factor(attrition$WorkLifeBalance)  
```
####### *(comment: Joblevel description was missing from the original dataset, so we only have the category number).

#### View a Summary of the data 
```{r, include=TRUE, echo=TRUE, fig.width=10}
summary(attrition)
# attach for ease of running code 
attach(attrition)
```

#### Some initial things to note here: 
* Only 16% attrition in the data 
* BusinessTravel, OverTime, JobInvolvement, PerformanceRating, seem to be unbalanced. All other Variables seem balanced. 
* YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager may be skewed

#### Need logistic regression

#### *** The next several steps are the model fitting steps 
#### Model fitting (Steps 1 thru 8) 

##### 1. Split the data into train and test data 
###### We want to be able to create an appropriate model the test that model on a subset of the data. We will create a "training" dataset and a "test" one. 
```{r, include=TRUE, echo=TRUE, fig.width=10}
set.seed(123)
split <- sample.split(attrition$Attrition, SplitRatio = 0.80)
#get training and test data
data.train <- subset(attrition, split == TRUE)
data.test <- subset(attrition, split == FALSE)
```

##### 2. Define null + full models for the training data
######  We need a NULL model (with no variables) to compare our training model to. We will compare the Deviance and AIC of the NULL Model vs the training model. We are looking for significant drop in both of these stats as a validation that the model is working. 
```{r, include=TRUE, echo=TRUE, fig.width=10}
options(width = 300) ## print 
null.model.train<-glm(formula=Attrition~1, family = binomial(link="logit"), data=data.train)
full.model.train<-glm(formula=Attrition~., family = binomial(link="logit"), data=data.train)
summary(full.model.train)
```

##### There are a few things to note here:
* We can see the far-right column with asterisks representing the lowest p-values. Since this is the whole model, and it has not been "fitted," the values here represent individual variables, that, on their own, they each appear to have some significance and explain Attrition.  
* The "Estimate" column gies us the Log Odds (logit scale). This will back-transformed in a later step to actual Odds-Ratio, which is the probability that a covariate value influences Attrition in terms of Odds.   
* The Null Deviance (1040.18) is a measure of lack of fit of the model without any variables. 
* The Residual Deviance (633.03) is a measure of lack of fit of the model with all variables included.  
* AIC (761.03) is a measure of how well the model is fitting. We want to keep this as low as possible. 

######  Sorting Model by p-value descending to view important covariates
```{r, include=TRUE, echo=TRUE, fig.width=10}
#sort summary by p-value descending
tempDF<-as.data.frame(summary(full.model.train)$coefficients)
tempDF[order(tempDF$`Pr(>|z|)`),]
```
###### We can see the top variables, that, on their own, seem to have an effect in Attrition. 

##### 3. Run ANOVA 
###### We want to run an ANOVA To further understand the impact of each variable into the model. 
```{r, include=TRUE, echo=TRUE, fig.width=10}
anova(full.model.train, test="Chisq")
```
###### If we take a look at the Residual Deviance column of ANOVA above, we can get a feel for how much impact a single variable has in the model. The ANOVA calculates the Residual Deviance (lack of fit) that each variable contributes to. Each variable is introduced in order of appearance in dataset (column-wise). For instance, starting from Null model (1040.18), when the first variable, "BusinessTravel" is introduced, there is a drop of Deviance of 18 points. Later variables may contribute more or less to the overall Deviance. The variables that drop the Deviance to the greatest degree, are the ones that may be heavily contributing to Attrition. 


##### 4.  Fitting the model using STEP 
###### We will use STEP as selection procedure (direction="both"), which starts with a NULL model and adding one variable at the time, validating the impact. If the variable is not meaningful it will be discarded. This process is repeated until we go thru all the variables. (Long output ommited, only last few lines shown here for brevity).
```{r, include=FALSE, echo=FALSE, fig.width=10}
step(null.model.train, scope=list(upper=full.model.train), direction="both", test="Chisq", data=data.train)
```
```{r, include=TRUE, echo=FALSE, fig.width=10}
cat('step(null.model.train, scope=list(upper=full.model.train), direction="both", test="Chisq", data=data.train)')
cat("
. . . 
< many steps later >
. . . 
                           Df Deviance    AIC    LRT  Pr(>Chi)    
<none>                          653.05 745.05                     
...
- DistanceFromHome          1   668.28 758.28 15.224 9.547e-05 ***
- JobSatisfaction           3   672.87 758.87 19.820 0.0001850 ***
- JobInvolvement            3   673.25 759.25 20.195 0.0001546 ***
- NumCompaniesWorked        1   672.21 762.21 19.153 1.207e-05 ***
- BusinessTravel            2   675.76 763.76 22.705 1.174e-05 ***
- EnvironmentSatisfaction   3   683.97 769.97 30.921 8.831e-07 ***
- JobLevel                  4   686.83 770.83 33.779 8.271e-07 ***
- JobRole                   8   704.21 780.21 51.153 2.452e-08 ***
- OverTime                  1   743.18 833.18 90.130 < 2.2e-16 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
. . .
Degrees of Freedom: 1175 Total (i.e. Null);  1130 Residual
Null Deviance:       1040 
Residual Deviance: 649.2     AIC: 741.2 
")
```
###### The output of the Step procedure above, shows most significant variables (in descending order). We will choose these top 9 (lowest p-value) variables for our Final Training Model. 

##### 5.  Create Final Training Model with top variables (lowest p-value). 
```{r, include=TRUE, echo=TRUE, fig.width=10}
final.model.train<-glm(formula=Attrition~OverTime+JobRole+JobLevel+EnvironmentSatisfaction+BusinessTravel+NumCompaniesWorked+JobInvolvement+JobSatisfaction+DistanceFromHome, family = binomial(link="logit"), data=data.train)
summary(final.model.train)
```

##### Sort the model by p-value 
```{r, include=TRUE, echo=TRUE, fig.width=10}
#sort summary by p-value descending
tempDF<-as.data.frame(summary(final.model.train)$coefficients)
tempDF[order(tempDF$`Pr(>|z|)`),]
```

##### 6.  Test Final Training Model Against Test data 
###### We are looking for a measure of how accurate this Training Model is against a subset of the data. 
```{r, include=TRUE, echo=TRUE, fig.width=10}
predict.results <- predict(final.model.train,newdata=subset(data.test,select=c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32)),type='response')
predict.results <- ifelse(predict.results > 0.5,1,0)
misClasificError <- mean(predict.results)
print(paste('Accuracy',1-misClasificError))
```
###### We see a 91% accuracy of the model against test data, so we are good. 

##### 7. Run Final model against all data. 
###### Now we ned to run the Final model against all the data. 
```{r, include=TRUE, echo=TRUE, fig.width=10}
final.model.alldata<-glm(formula=Attrition~OverTime+JobRole+JobLevel+EnvironmentSatisfaction+BusinessTravel+NumCompaniesWorked+JobInvolvement+JobSatisfaction+DistanceFromHome, family = binomial(link="logit"), data=attrition)
summary(final.model.alldata)
```

##### 8. Obtain Odd Ratio of covariences 
###### The Estimate in the final model is given in terms of Log Odds. Need to invert to obtain actual Odd Ratios of the top variables and sort output
```{r, include=FALSE, echo=TRUE, fig.width=10}
sort(exp(final.model.alldata$coefficients[-1]),decreasing=T) 
```
###### Looking at the odd ratio table below, we can see that the Sales Representatives category has the highest odds of causing Attrition, this is, Sales representatives are 8 times more likely to have Attrition than any other category. The second highest covariate is Working Overtime. people who work overtime are 5 times more likely to have Atrition. People who travel frequently are 4x more likely to leave. Also Job Roles like Lab technician and HR are more likely to leave, based on the data.   

Covarience/Value	|	Odds Ratio
--------------------------------	|	----------
JobRoleSales Representative	|	8.3669402
OverTimeYes	|	5.3426139
BusinessTravelTravel_Frequently	|	4.7865786
JobRoleLaboratory Technician	|	3.6904708
JobRoleHuman Resources	|	3.4551899
JobRoleSales Executive	|	3.1500722
BusinessTravelTravel_Rarely	|	2.2161456
JobRoleResearch Scientist	|	1.7691884
JobRoleManager	|	1.5547907
NumCompaniesWorked	|	1.119052
DistanceFromHome	|	1.0362319 

###### REFERENCES 
Czado, C. (1997). On selecting parametric link transformation families in generalized linear models. J. Statist. Plann. Inference 61, 125–139.
Hastie, T. and R. Tibshirani (1999). Generalized additive models (2nd edition). London: Chapman & Hall.
Author, A. A. (Year of publication). Title of work: Capital letter also for subtitle. Location: Publisher.
Ramsey, Fred (2013). The Statistical Sleuth. Chapters 20-23. 
Lecture 20 - Logistic Regression. Retrieved from https://www2.stat.duke.edu/courses/Spring13/sta102.001/Lec/Lec20.pdf
Multivariate Logistic Regression. Retrieved from http://www.med.mcgill.ca/epidemiology/joseph/courses/EPIB-621/logistic2.pdf
Multiple logistic regression. Handbook of Biological Statistics. Retrieved from http://www.biostathandbook.com/multiplelogistic.html
Binary Response and Logistic Regression Analysis. Retrieved from http://www.stat.wisc.edu/~mchung/teaching/MIA/reading/GLM.logistic.Rpackage.pdf
An R Companion for the Handbook of Biological Statistics. Retrieved from https://rcompanion.org/rcompanion/e_07.html
Simple Guide to Logistic Regression in R. Retrieved from https://www.analyticsvidhya.com/blog/2015/11/beginners-guide-on-logistic-regression-in-r/
How to Perform a Logistic Regression in R. Retrieved from https://datascienceplus.com/perform-logistic-regression-in-r/
Cookbook for R. Logistic regression. Retrieved from http://www.cookbook-r.com/Statistical_analysis/Logistic_regression/ 

### Job Roles

####<font color="midnightblue">Job Roles</font>  
*Note: All analysis has been performed on employees only (No attrition participants)*

####Key Findings

**1) Manager Job Role:**  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;a) Only job role that exists in all departments.  All other job roles are unique to a department.  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;b) Have the longest average duration at the company  
**2) Managers and Directors**  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;a) Tend to be older than people in other job roles  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;b) Have higher Job Levels  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;c) Have  higher Monthly Income  
**3) Manufacturing Director**  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;a) The only Job Role where Males do not outnumber Females  
**4) Human Resource Manager**  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;a) is the only Role which does not have 'Low' level of job involvement  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;b) No PhDs (characteristic in common with Sale Reps)  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;c) All HR Managers have some College level education (a characteristic unique to this Job Role)  
**5) Sales Reps and Reseach Scientists**  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;a) have the lowest Job Levels  
**6) Sales Reps**  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;a) is the only Job Role which has no employees with NumWorked = 9!!  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;b) have the lowest average duration at the company  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;c) change managers most frequently, while directors and managers tend to stay with the same manager for a longer time  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;d) No PhDs (characteristic in common with HR Managers)  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;e) the highest percentage of employees with college-level education (25%)  


####<font color="midnightblue">Departments</font>  
```{r, fig.width=10}

ggplot(data=as.data.frame(table(noattrit$Department,noattrit$JobRole,dnn=list("Department","JobRole"))), aes(x= reorder(JobRole,Freq),y=Freq, fill = Department)) +
geom_bar(stat="identity") + coord_flip() + ggtitle("Number of Employees per Job Role by Department") + xlab("") + ylab("")+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4")))
```

**Linear Regression Modeling (See Experiment Tab) suggests Education Field, Age, Job Involvement, Gender, Job Level, Monthly Income, Number of Companies Worked, Years At Company and Years with Current Manager are different between of Job Roles**  

####<font color="midnightblue">Education</font>

```{r, fig.width=10}
ggplot(data=as.data.frame(table(noattrit$JobRole,noattrit$EducationField,dnn=list("JobRole","EducationField"))), aes(x= reorder(JobRole,Freq),y=Freq, fill = EducationField)) +
geom_bar(stat="identity",position="fill") + coord_flip() + ggtitle("Percent of Employees per Job Role by Education Field") + xlab("") + ylab("Percent of Employees")+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "midnightblue", "mistyrose4", "lightcyan3")))
```

*Though not identified as a first order differentiator by linear regression, Education is an interesting parameter in regards to Job Role*  

```{r, fig.width=10}
ggplot(data=as.data.frame(table(noattrit$JobRole,noattrit$Education,dnn=list("JobRole","Education"))), aes(x= reorder(JobRole,Freq),y=Freq, fill = Education)) +
geom_bar(stat="identity",position="fill") + coord_flip() + ggtitle("Education by Job Role \nNo PhD Sales Reps nor HR Managers \nAll HR Managers Have Some College") + xlab("") + ylab("Percent of Employees")+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "mistyrose4", "lightcyan3")))
```

##### Observations
**All Departments have a mix of educational fields**  
**People with Education in Marketing only work in the Sales Department**  
**People with Education in Human Resources work in the Human Resources Department**  
**R&D is predominantly composed of people with Life Sciences and Medical Education**  
**No PhD Sales Reps nor HR Managers**  
**All HR Managers Have Some College (unique from all other job roles)**  
**Sales Reps have the highest percentage of employees with college-level education (25%)** 

####<font color="midnightblue">Age</font>  
```{r, fig.width=10}
noattrit$JobRole <-with(noattrit, ifelse(noattrit$JobRole=="Manager", paste(noattrit$Department,noattrit$JobRole), as.character(noattrit$JobRole)))
ggplot(data=noattrit,aes(x= JobRole,y=Age)) +
geom_boxplot() + coord_flip() + ggtitle("Age by Job Role") + xlab("")+ylab("Age")+theme(plot.title=element_text(hjust=0.5))
#Descripitve Statistics of Employee Ages
  summary(noattrit$Age)
  results <- lm(noattrit$JRCode~noattrit$Age)
summary(results)
```

#####Observations
**1. Managers and Directors tend be older than people in other roles**  
**2. The oldest employee is a Sales Executive (60 years old)**  
**3. The youngest employee is a Laboratory Technician (18 years old)**  

####<font color="midnightblue">Gender</font>  
```{r, fig.width=10}
ggplot(data=as.data.frame(table(noattrit$JobRole,noattrit$Gender,dnn=list("JobRole","Gender"))), aes(x= reorder(JobRole,Freq),y=Freq, fill = Gender)) +
geom_bar(stat="identity",position="fill") + coord_flip() + ggtitle("Males OutNumber Females in Most Job Roles \n (Delta is not statiscally significant)") + xlab("") + ylab("Percent of Employees")+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4"))) + geom_hline(data=noattrit, aes(yintercept = .50), colour="black") 
```



####<font color="midnightblue">Compensation</font>  
```{r, fig.width=10}
ggplot(data=noattrit,aes(x= JobRole,y=MonthlyIncome,group=JobRole)) + geom_boxplot() + coord_flip() + ggtitle("Managers and Directors Have Higher Compensation \n (p-value < .00001)") + xlab("")+ylab("Monthly Income")+theme(plot.title=element_text(hjust=0.5))

```

####<font color="midnightblue">Job Levels</font>
```{r, fig.width=10}
 ggplot(data=as.data.frame(table(noattrit$JobRole,noattrit$JobLevel,dnn=list("JobRole","JobLevel"))), aes(x= reorder(JobRole,Freq),y=Freq, fill = JobLevel)) + geom_bar(stat="identity",position="fill") + coord_flip() + ggtitle("Managers & Directors Tend to Have Higher Job Levels") + xlab("") + ylab("Percent of Employees")+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "mistyrose4", "lightcyan3")))

results <- lm(noattrit$JRCode~noattrit$JobLevel)
summary(results)

```

#####Observations:
**1. Managers and Directors have higher Job Levels.**  
**2. Sales Reps and Research Scientists have the lowest Job Levels.**  

####<font color="midnightblue">Job Involvement</font>  
```{r, fig.width=10}
ggplot(data=as.data.frame(table(attrit$JobRole,attrit$JobInvolvement,dnn=list("JobRole","JobInvolvement"))), aes(x= reorder(JobRole,Freq),y=Freq, fill = JobInvolvement)) +
geom_bar(stat="identity",position="fill") + coord_flip() + ggtitle(" HR Managers are the only Role \nwithout Low Job Involvement") + xlab("") + ylab("Percent of Employees") +theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "mistyrose4")))
```

####<font color="midnightblue">Work Experience</font>  
```{r, fig.width=10}
ggplot(data=noattrit,aes(x= JobRole,y=noattrit$NumCompaniesWorked)) +
geom_boxplot() + coord_flip() + ggtitle("Many Employees have Experience at Several Companies") + xlab("")+ylab("Number of Companies") +theme(plot.title=element_text(hjust=0.5))
```

#####Observations:
**Sales Rep is the only Job Role which has no employees with NumWorked = 9!!**  

####<font color="midnightblue">Years at Company</font>  
```{r, fig.width=10}
ggplot(data=noattrit,aes(x= JobRole,y=noattrit$YearsAtCompany)) + geom_boxplot() + coord_flip() + ggtitle("Significant Differences between Tenure by Job Role \n(p-value=.0008)") + xlab("")+ylab("Number of Years at Company") +theme(plot.title=element_text(hjust=0.5))
  
results <- lm(noattrit$JRCode~noattrit$YearsAtCompany)
summary(results)
  
```

#####Observations
**The Job Role "Sales Reps" has the lowest average duration at the company.**  
**The Job Roles of Managers have the longest average duration at the company.**  


```{r, fig.width=10}
ggplot(data=noattrit,aes(x= JobRole,y=YearsWithCurrManager)) + geom_boxplot() + coord_flip() + ggtitle("Years With Current Manager Vary By Job Role \n (p-value = .0425)") + xlab("")+ylab("Number of Years")+theme(plot.title=element_text(hjust=0.5))
```

#####Observations:
**Sales Reps change managers most frequently, while directors and managers tend to stay with the same manager for a longer time.**  


### Company Specifics
#### <font color="midnightblue">Company Specifics</font>   

##### <font color="midnightblue">Key Findings</font>
**67% of the Employees in the Company in R&D**  
**The average age of employees in 37.56 years, with the employees in HR average age being slightly higher**  
**There are 46% more Male employees than Female employees.  HR has the largest ratio of Males to Female at 2.6:1**  
**Job Level is strongly correlated to Montly Income. High Job Level = Higher Monthly Incomes**  
**Human Resource Managers are the only Role which does not have 'Low' level of job involvement**  
**50% of Employees have been with the company <= 5 years**  
**People with Marketing education only work in Sales, and people with HR education only work in HR.  All departments have a mix of Educational Backgrounds**  
**Everyone receives a Performance Rating of Excellent or Outstanding**  
**Work Experience correlates to Job Level and therefore Monthly Income**  
**Monthly Income is based on Work Experience, not Performance. (Job Level is correlated to Total Time Worked and Time At Company. Monthly Income is correlated to Job Level and Total Time Worked.  All Performance Ratings are Excellent and Outstanding.)**  
**The longer you work for a manager the longer the time between promotions**  



##### <font color="midnightblue">Departments</font>  
```{r, fig.width=10}

DepartmentDF <- as.data.frame(table(noattrit$Department))
colnames(DepartmentDF) <- c("Department","DepartmentCount")

DepartmentDF$CoDepartmntPercentage <-paste(100*as.numeric(format(DepartmentDF$DepartmentCount/colSums(DepartmentDF[2]),digits=1)),"%")
DepartmentDF$CoDepartmntPercentage <-paste(100*as.numeric(format(DepartmentDF$DepartmentCount/colSums(DepartmentDF[2]),digits=1)),"%")

 print.data.frame(DepartmentDF[with(DepartmentDF, order(-DepartmentCount)),],row.names=FALSE)
 
ggplot(data=as.data.frame(table(noattrit$Department,dnn=list("Department"))), aes(x= reorder(Department,Freq),y=Freq)) +
geom_bar(stat="identity", fill="steelblue4") + coord_flip() + ggtitle("Department Sizes \n(no attrition)") + xlab("") + ylab("Number of Employees")+theme(plot.title=element_text(hjust=0.5))
```

#####<font color="midnightblue">Age Distribution</font>

```{r, fig.width=10}
options(width = 150)
ggplot(data=noattrit, aes(x=noattrit$Age,fill=noattrit$Department)) + geom_histogram()+theme(plot.title=element_text(hjust=0.5))+xlab("Age")+ylab("Number of Employees")+ggtitle("Age Distribution by Department")+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4")))
cat("Summary of Age Distribution for the Company")
summary(noattrit$Age)
cat("Summary of Age Distributions by Department")
aggdata <-aggregate(noattrit$Age, by=list(noattrit$Department), 
FUN=summary, na.rm=TRUE)
print(aggdata)

```
#####<font color="midnightblue">Gender Distribution by Department</font>
```{r, fig.width=10}
ggplot(data=as.data.frame(table(noattrit$Department,noattrit$Gender,dnn=list("Department","Gender"))), aes(x= reorder(Department,Freq),y=Freq, fill = Gender)) + geom_bar(stat="identity",position="fill") + coord_flip() + ggtitle("Gender by Department") + xlab("") + ylab("Percent of Employees") + geom_hline(data=noattrit, aes(yintercept = .50), colour="black") +theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4")))
ggplot(data=noattrit, aes(x=noattrit$Gender,fill=Gender)) + geom_bar() + ggtitle("There are 46% more Males than Females in the Company \n(p-value=.04)") + xlab("Gender")+ylab("Number of Employees")+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4")))

#Recode Department to a numeric for t-test analysis
noattrit$DeptCode[noattrit$Department=="Research & Development"] <- 1L
noattrit$DeptCode[noattrit$Department=="Sales"] <- 2L
noattrit$DeptCode[noattrit$Department=="Human Resources"] <- 3L
t.test(table(noattrit$DeptCode, noattrit$Gender))
 

```

```{r, fig.width=10}
ggplot(data=noattrit,aes(x=noattrit$JobLevel, y=noattrit$MonthlyIncome)) + geom_point()  + ggtitle("Monthly Income is correlated to Job Level \n(p-value=<.0001)") + ylab("Monthly Income") +xlab("Job Level") + geom_smooth(method = 'lm', se = FALSE, color="midnightblue")+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4")))
results <- lm(noattrit$MonthlyIncome~noattrit$JobLevel)
summary(results)

```
#####<font color="midnightblue">Job Involvement</font>
```{r, fig.width=10}
 ggplot(data=as.data.frame(table(noattrit$Department,noattrit$JobInvolvement,dnn=list("Department","JobInvolvement"))), aes(x= reorder(Department,Freq),y=Freq, fill = JobInvolvement)) + geom_bar(stat="identity",position="fill") + coord_flip() + ggtitle("4.4% of the Company has a Low Level of Job Involvement") + xlab("") + ylab("") + geom_hline(data=noattrit, aes(yintercept = .9554), colour="black")+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "mistyrose4")))

t.test(table(noattrit$DeptCode, noattrit$JobInvolvement))
```

####<font color="midnightblue">Number of Years with Current Manager</font>


```{r, fig.width=10}
ggplot(data=noattrit, aes(x=factor(noattrit$YearsWithCurrManager))) + geom_histogram(stat="count",position = "dodge", fill="steelblue4") +xlab("Years")+ylab("Number of Employees") + scale_x_discrete(name='Years With Current Manager') + ggtitle("Years With Current Manager")+theme(plot.title=element_text(hjust=0.5))
```

#####Observations:
**Multiple Modal Distribution**  

#####<font color="midnightblue">Education</font>

```{r, fig.width=10}
cat("Percentage of Employees with No College-Level Education:",paste(100*as.numeric(format(nrow(as.data.frame(noattrit[( noattrit$Education == 1),]))/1233,digits=1)),"%"))

ggplot(data=noattrit, aes(x=factor(noattrit$Education))) + geom_histogram(stat="count",position = "dodge", fill="steelblue4") +xlab(" ") + scale_x_discrete(name='Education') + ggtitle("Education Level")+theme(plot.title=element_text(hjust=0.5))
```



####Compensation is based on Work Experience not Performance

#####<font color="midnightblue">Job Level vs. Total Working Years</font>
```{r, fig.width=10}
ggplot(data=noattrit,aes(x= JobLevel,y=noattrit$TotalWorkingYears, group=JobLevel)) + geom_boxplot() + coord_flip() + ggtitle("Work Experience Correlates to Job Level") + xlab("Job Level")+ylab("Years")+theme(plot.title=element_text(hjust=0.5))

```

#####<font color="midnightblue">Working Years vs. Monthly Income</font>
```{r, fig.width=10}
ggplot(data=noattrit,aes(x=noattrit$TotalWorkingYears,
y=noattrit$MonthlyIncome,col=factor(noattrit$JobLevel))) +
geom_point()  + ggtitle("Total Working Years is Correlated to Job Level \n(p-value <.00001)") + xlab("Working Years")+ylab("Monthly Income")+theme(plot.title=element_text(hjust=0.5))+scale_color_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "mistyrose4", "lightcyan3")))

results <- lm(noattrit$JobLevel~noattrit$TotalWorkingYears)
summary(results)

 
```

#####<font color="midnightblue">Total Time at Company vs. Monthly Income</font>

```{r, fig.width=10}
ggplot(data=noattrit,aes(x=noattrit$YearsAtCompany, y=noattrit$MonthlyIncome,col=factor(noattrit$JobLevel))) +
geom_point()  + ggtitle("Years At Company vs Monthly Income by JobLevel") + xlab("Years at Company")+ylab("Monthly Income")+theme(plot.title=element_text(hjust=0.5))+scale_color_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "mistyrose4", "lightcyan3")))

 results <- lm(noattrit$MonthlyIncome ~ noattrit$JobLevel + noattrit$TotalWorkingYears + noattrit$YearsAtCompany)
 summary(results)

```


##### Performance Rating
```{r, fig.width=10}
ggplot(data=noattrit, aes(x=factor(noattrit$PerformanceRating))) + geom_histogram(stat="count",position = "dodge", fill="steelblue4") +xlab(" ") + scale_x_discrete(name='Performance Rating') + ggtitle("Everyone Receives a Performance Rating of Excellent or Outstanding")+theme(plot.title=element_text(hjust=0.5))

ggplot(data=noattrit,aes(x= PerformanceRating, y=noattrit$MonthlyIncome, group=PerformanceRating)) + geom_boxplot() + ggtitle("Monthly Income by Performance Rating") + xlab("Performance Rating")+ylab("Monthly Income")+theme(plot.title=element_text(hjust=0.5))
```

#####<font color="midnightblue">Years Since Promotion</font>
```{r, fig.width=10}
ggplot(data=noattrit,aes(x=noattrit$YearsWithCurrManager, y=noattrit$YearsSinceLastPromotion)) +
geom_point()  + ggtitle("The Longer You Stay with a Manager the Longer Time Between Promotions") + xlab("Years With Current Manager")+ylab("Years Since Last Promotion") + stat_smooth(method="lm", se=FALSE, color="midnightblue")+theme(plot.title=element_text(hjust=0.5))+scale_color_manual(values=alpha(c("steelblue4")))
results <- lm(noattrit$YearsSinceLastPromotion ~ noattrit$YearsWithCurrManager)
summary(results)

ggplot(data=noattrit,aes(x= JobLevel,y=noattrit$YearsSinceLastPromotion, group=JobLevel)) + geom_boxplot() + coord_flip() + ggtitle("Time Since Last Promotion Increases as Job Level Increases") + xlab("Job Level")+ylab("Years")+theme(plot.title=element_text(hjust=0.5))

```

####<font color="midnightblue">Percent Salary Hike</font>

```{r, fig.width=10}

ggplot(data=noattrit,aes(x= JobLevel,y=noattrit$PercentSalaryHike, group=JobLevel)) + geom_boxplot() + coord_flip() + ggtitle("All Job Levels Received About 15% Average Salary Increase \nwith roughly similar distributions") + xlab("Job Level")+ylab("% Salary Increase")+theme(plot.title=element_text(hjust=0.5))

```




### Job Satisfaction
####<font color="midnightblue">Job Satisfaction</font>
#####<font color="midnightblue">Key Findings</font>
**Job Satisfaction Increases as Hourly Rate Decreases**  
**PhDs have the highest satisfaction**  
**People with Bachelor Degrees are the most dissatified**  
**Satisfaction in consistent by Gender**  
**Satisfaction varies by Job Role with R&D Manager having the lowest satisfaction and Sales Reps and HR Managers having the highest**  

#####<font color="midnightblue">Hourly Rate</font>
**Suggested by Linear Regression Modeling (See Experimentation Tab) as weak factor for Job Satisfaction (other potential factors were identified but did not pass on review)**  

```{r, fig.width=10}
ggplot(data=noattrit,aes(x= JobSatisfaction,y=noattrit$HourlyRate, group=JobSatisfaction)) + geom_boxplot() + coord_flip() + ggtitle("Job Satisfaction Inreases as Hourly Rate Decreases \np-value=.05") + xlab("Job Satisfaction")+ylab("Hourly Rate")+theme(plot.title=element_text(hjust=0.5))

```

**Visual inspection of the Top 10 common demographics suggests exploring Department, Education Field, Gender and Job Role as factors related to Job Satisfaction**  

```{r}
#recode JobSatisfaction for analysis
noattrit$JSCode[noattrit$JobSatisfaction==1] <- "Low"
noattrit$JSCode[noattrit$JobSatisfaction==2 |noattrit$JobSatisfaction==3 | noattrit$JobSatisfaction==4] <- "High"
noattrit$JSCodeNum[noattrit$JobSatisfaction==2 |noattrit$JobSatisfaction==3 | noattrit$JobSatisfaction==4] <- 2L
noattrit$JSCodeNum[noattrit$JobSatisfaction==1] <- 1L

```

####<font color="midnightblue">Education</font>
```{r, fig.width=10}
ggplot(data=as.data.frame(table(noattrit$Education,noattrit$JobSatisfaction,dnn=list("Education","JobSatisfaction"))), aes(x=Education,y=Freq, fill = JobSatisfaction)) + geom_bar(stat="identity",position="fill") + ggtitle("PhDs Have the Highest Job Satisfaction \n People with Bachelor Degrees Have the Lowest Job Satisfaction \n(p-value=.01263)") + xlab("Education Level") + ylab("Percent of Employees") + geom_abline(slope=0, intercept=0.5,  col = "black",lty=2) + coord_flip()+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "mistyrose4")))

ggplot(data=as.data.frame(table(noattrit$Education,noattrit$JSCode,dnn=list("Education","JobSatisfaction"))), aes(x=Education,y=Freq, fill = JobSatisfaction)) + geom_bar(stat="identity",position="fill") + ggtitle("Job Satisfaction Varies by Education") + xlab("Education Level") + ylab("Percent of Employees") + coord_flip()+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4")))

t.test(table(noattrit$JSCodeNum, noattrit$Education))
```



####<font color="midnightblue">Gender</font>
```{r, fig.width=10}
ggplot(data=as.data.frame(table(noattrit$Gender,noattrit$JobSatisfaction,dnn=list("Gender","JobSatisfaction"))), aes(x=Gender,y=Freq, fill = JobSatisfaction)) +
geom_bar(stat="identity",position="fill") + ggtitle("Job Satisfaction is Consistent Across Genders") + xlab("Gender") + ylab("Percent of Employees") + geom_abline(slope=0, intercept=0.5,  col = "black",lty=2)+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "mistyrose4")))
```


####<font color="midnightblue">Job Role</font>

```{r, fig.width=10}
ggplot(data=as.data.frame(table(noattrit$JobRole,noattrit$JobSatisfaction,dnn=list("JobRole","JobSatisfaction"))), aes(x=JobRole,y=Freq, fill = JobSatisfaction)) + geom_bar(stat="identity",position="fill") + ggtitle("Job Satisfaction is Varies by Job Roles \nTop Dissatisfied Roles:R&D Director  \nMost Satsifed: Sales Reps and HR Manager") + xlab("") + ylab("Percent of Employees") + geom_abline(slope=0, intercept=0.9,  col = "black",lty=2) + coord_flip()+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "mistyrose4")))

t.test(table(noattrit$JSCodeNum, noattrit$JobRole))
```


### Interesting Findings

####<font color="midnightblue">Interesting Findings</font>
#####<font color="midnightblue">Key Findings</font>

**NumCompaniesWorked has evidence of ambiguity in the data and is perhaps miscoded or inaccurate.  (See discussion below)**  
**No relationship between the difference compensation rates.  Would have expected a multiplicative relationship between DailyRate and Hourly Rate, or Daily Rate and Monthly Rate, etc.**  




#####Observatons
**Strong Correlation between observation incoming order and employee number**  


#####<font color="midnightblue">Number of Companies Worked</font>
```{r, fig.width=10}
summary(noattrit$NumCompaniesWorked)

ggplot(data=noattrit, aes(x=factor(noattrit$NumCompaniesWorked))) + geom_histogram(stat="count",position = "dodge", fill="steelblue4") +xlab("Number CompaniesWorked") + scale_x_discrete(name='Number of Companies') +ggtitle("Number of Companies Worked")+ylab("Number of Employees")+theme(plot.title=element_text(hjust=0.5))
```

#####Observations
**There maybe confusion in the coding of this variable.  What does 0 mean?  Since these are employees at this company everyone has worked at least one company.  Or maybe the variable is to be interpreted as the number of companies worked other than this one.  Seems unlikely as the most frequent value are**  


#### <font color="midnightblue">Compensation Rates - No correlations</font>

##### <font color="midnightblue">Monthly Income vs. Monthly Rate</font>

```{r, fig.width=10}
ggplot(data=noattrit,aes(x=noattrit$MonthlyRate, y=noattrit$MonthlyIncome)) +
geom_point()  + ggtitle(" No Correlation between Monthly Income vs Monthly Rate \n p-value=.261") + xlab("Monthly Rate")+ylab("Monthly Income") + stat_smooth(method="lm", se=FALSE, color="midnightblue")+theme(plot.title=element_text(hjust=0.5))

result <- lm(noattrit$MonthlyIncome ~ noattrit$MonthlyRate)
summary(result)

ggplot(data=noattrit,aes(x=noattrit$DailyRate, y=noattrit$MonthlyIncome)) +
geom_point()  + ggtitle(" No Correlation between Monthly Income vs Daily Rate") + xlab("Daily Rate")+ylab("Monthly Income") + stat_smooth(method="lm", se=FALSE, color="midnightblue")+theme(plot.title=element_text(hjust=0.5))

ggplot(data=noattrit,aes(x=noattrit$DailyRate, y=noattrit$MonthlyRate)) +
geom_point()  + ggtitle(" No Correlation between Monthly Rate & Daily Rate") + xlab("Daily Rate")+ylab("Monthly Rate") + stat_smooth(method="lm", se=FALSE, color="midnightblue")+theme(plot.title=element_text(hjust=0.5))

ggplot(data=noattrit,aes(x=noattrit$HourlyRate, y=noattrit$DailyRate)) +
geom_point()  + ggtitle(" No Correlation between Hourly Rate & Daily Rate") + xlab("Hourly Rate")+ylab("Daily Rate") + stat_smooth(method="lm", se=FALSE, color="midnightblue")+theme(plot.title=element_text(hjust=0.5))

```
```{r, fig.width=10}
ggplot(data=noattrit,aes(x=noattrit$YearsAtCompany, y=noattrit$DailyRate,col=factor(noattrit$JobLevel))) +
geom_point()  + ggtitle("Daily Rate vs Years At Company by JobLevel") + xlab("Years")+ylab("Daily Rate")+theme(plot.title=element_text(hjust=0.5))+scale_color_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "mistyrose4", "lightcyan3")))
```

**The strong correlations found between TimeAtCompany & Total Working Time with Monthly Income are not evident with DailyRate, Hourly Rate nor Monthly Rate.**  
**Attempted to determine relationships using log-log, linear-log, log-linear transformations and multiplicative models.**   





### Experimentation
####<font color="midnightblue">Experimentation</font>
1.Employee Number - a recoded start date?....No  
2.Linear Regression Modeling to id first order variables to differentiate factor variables  


**Employee Number is strongly correlated to the index number of the incoming data.  Perhaps the employee number is a time ordered variable related to employee start date?**  

```{r, fig.width=10}

ggplot(data=noattrit,aes(x=as.integer(rownames(noattrit)), y=noattrit$EmployeeNumber)) +
geom_point() + ggtitle("Index Number vs Employee Number") + ylab("Employee Number")+ xlab("Index Number")+theme(plot.title=element_text(hjust=0.5))

ggplot(data=noattrit,aes(x=noattrit$EmployeeNumber, y=noattrit$YearsAtCompany)) +
geom_point()  + ggtitle("Years At Company vs Employee \nEmployee Number is not related to start date") + xlab("Employee Number")+ylab("Years")+theme(plot.title=element_text(hjust=0.5))
```

**Used Linear Regression Modeling to determine variables of significance between Job Roles**  
**Important first order terms in the model are: EducationField, Age, JobLevel, MonthlyIncome, Gender, JobInvolvement, NumCompaniesWorked, YearsAtCompany, YearsWithCurrManager**  
```{r, fig.width=10}

options(width = 150)

#Run a linear regression model using all variables to id the important first order terms
jr.lm <- lm(noattrit$JRCode ~ noattrit$BusinessTravel+ noattrit$EducationField+ noattrit$Gender+ noattrit$OverTime+ noattrit$MaritalStatus+ noattrit$Age+ noattrit$DailyRate+ 
noattrit$DistanceFromHome+ noattrit$Education+ noattrit$EmployeeNumber+noattrit$EnvironmentSatisfaction+ noattrit$HourlyRate+ noattrit$JobInvolvement+ noattrit$JobLevel+ noattrit$JobSatisfaction+
 noattrit$MonthlyIncome+ noattrit$MonthlyRate+ noattrit$NumCompaniesWorked+ noattrit$PercentSalaryHike+ noattrit$PerformanceRating+ noattrit$RelationshipSatisfaction+ noattrit$StockOptionLevel+ 
noattrit$TotalWorkingYears+ noattrit$TrainingTimesLastYear+ noattrit$WorkLifeBalance+ noattrit$YearsAtCompany+ noattrit$YearsInCurrentRole+ noattrit$YearsSinceLastPromotion+ noattrit$YearsWithCurrManager)
summary(jr.lm)

#Check revised model fit
jr.lm <- lm(noattrit$JRCode ~noattrit$EducationField+ noattrit$Age + noattrit$JobLevel+ noattrit$MonthlyIncome+ noattrit$Gender+ noattrit$JobInvolvement+ noattrit$NumCompaniesWorked+ noattrit$YearsAtCompany+ noattrit$YearsWithCurrManager)
summary(jr.lm)

```
*The fit is not great, but perhaps sufficient enough to identify the important fit order terms to explore Job Roles*  

**Linear Regression Modeling for Job Satisfaction first order factors**  
```{r, fig.width=10}
js.lm <- lm(noattrit$JobSatisfaction ~ noattrit$BusinessTravel+ noattrit$EducationField+ noattrit$Gender+ noattrit$OverTime+ noattrit$MaritalStatus+ noattrit$Age+ noattrit$DailyRate+ noattrit$DistanceFromHome+ noattrit$Education+ noattrit$EmployeeNumber+ noattrit$EnvironmentSatisfaction+ noattrit$HourlyRate+ noattrit$JobInvolvement+ noattrit$JobLevel+ noattrit$JobRole+ noattrit$MonthlyIncome+ noattrit$MonthlyRate+ noattrit$NumCompaniesWorked+ noattrit$PercentSalaryHike+ noattrit$PerformanceRating+ noattrit$RelationshipSatisfaction+ noattrit$StockOptionLevel+ noattrit$TotalWorkingYears+ noattrit$TrainingTimesLastYear+ noattrit$WorkLifeBalance+ noattrit$YearsAtCompany+ noattrit$YearsInCurrentRole+ noattrit$YearsSinceLastPromotion+ noattrit$YearsWithCurrManager)
summary(js.lm)


js.lm <- lm(noattrit$JobSatisfaction ~noattrit$OverTime+noattrit$MaritalStatus+noattrit$HourlyRate + noattrit$RelationshipSatisfaction)

summary(js.lm)

#####Note:  The fit is poor, but may provide directionality as to first order factors to assess.
```

**Linear Regression Modeling suggests that Overtime, Marital Status, Hourly Rate, and Relationship Satisfaction may be factors related to Job Satisfaction. This method didn't work as well as with understanding of Job Roles.  Only one of the four variables has a good correlation to Job Satisfaction, Hourly Rate.**  


#####<font color="midnightblue">Overtime</font>
```{r, fig.width=10}

ggplot(data=as.data.frame(table(attrit$JobSatisfaction,attrit$OverTime,dnn=list("JobSatisfaction","OverTime"))), aes(x= reorder(JobSatisfaction,Freq),y=Freq, fill = OverTime)) +
geom_bar(stat="identity",position="fill") + coord_flip() + ggtitle("Percent of People Who Have Over Time is Consistent for All Job Satisfation Levels") + xlab("") + ylab("Percent of Employees") + geom_abline(slope=0, intercept=0.5,  col = "black",lty=2)+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4")))

```

#####<font color="midnightblue">Marital Status</font>
```{r, fig.width=10}
ggplot(data=as.data.frame(table(attrit$JobSatisfaction,attrit$MaritalStatus,dnn=list("JobSatisfaction","MaritalStatus"))), aes(x= reorder(JobSatisfaction,Freq),y=Freq, fill = MaritalStatus)) +
geom_bar(stat="identity",position="fill") + coord_flip() + ggtitle("Job Satisfaction by Marital Status is Consistent") + xlab("Job Satisfaction Level") + ylab("Percent of Employees") + geom_abline(slope=0, intercept=0.5,  col = "black",lty=2)+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4")))

```

##### <font color="midnightblue">Relationship Satisfaction</font>
```{r, fig.width=10}
ggplot(data=as.data.frame(table(attrit$JobSatisfaction,attrit$RelationshipSatisfaction,dnn=list("JobSatisfaction","RelationshipSatisfaction"))), aes(x= reorder(JobSatisfaction,Freq),y=Freq, fill = factor(RelationshipSatisfaction))) +
geom_bar(stat="identity",position="fill") + coord_flip() + ggtitle("Job Satisfaction by Relationship Satisfaction is Consistent") + xlab("Job Satisfaction Level") + ylab("Percent of Employees") + geom_abline(slope=0, intercept=0.5,  col = "black",lty=2)+theme(plot.title=element_text(hjust=0.5))+scale_fill_manual(values=alpha(c("steelblue4", "thistle4", "paleturquoise4", "mistyrose4")))

```

**Find common demographics for people with high/lo job satisfaction**  
```{r}
JS.4 <- noattrit[noattrit$JobSatisfaction==4,]
JS.3 <- noattrit[noattrit$JobSatisfaction==3,]
highjs <- rbind(JS.3, JS.4)
demohighjs<- highjs[1:8]
demohighjs$Attrition <- NULL
demohighjs$BusinessTravel <- NULL
demohighjs$OverTime <- NULL
demohighjs$MaritalStatus <- NULL
countdemohighjs<- count(demohighjs)
head(countdemohighjs[order(countdemohighjs$freq,decreasing = TRUE),],10)

```
**Visual inspection of the Top 10 common demographics suggests exoloring Department, Education Field, Gender and Job Role as factors related to Job Satisfaction**  

==















